SeisAcoustic

### 2019-05-22 docs/latex/Acoustic_FWI/paper.tex
frequency domain gradient

We use stagger grid finite difference method to discretize equation \ref{eq2}, the particle $v_z$ is put on the stagger grid $[i+\frac{1}{2},j,it]$, horizontal particle velocity $v_x$ is put on the stagger grid $[i, j+\frac{1}{2}, k]$, and the pressure component is put on the grid $[i,j,it+\frac{1}{2}]$. Where $i,j,it$ are the grid index in vertical, horizontal, time directions, respectively. Figure \ref{f1} shows the locations of wave field components and model parameters, respectively.
\begin{figure}[htb] 
   \centering
   \includegraphics[width=0.6\columnwidth]{Fig/f1.pdf} 
   \caption{Stagger grid}
   \label{f1}
\end{figure}

@@

\subsection{${\bf p}_x$ time-stepping}
\begin{equation}
\frac{\partial p_x}{\partial t} + \gamma_x p_x = k \frac{\partial v_x}{\partial x} + f(t) 
\label{eq32}
\end{equation}
The finite difference form of this equation is given as
\begin{equation}
\begin{split}
&p_x[i,j,it+\frac{3}{2}] = \frac{2-dt \, \gamma_x[j]}{2+dt \, \gamma_x[j]} p_x[i,j,it+\frac{1}{2}]  \\
&+ \frac{2k[i,j]\, dt}{2+dt \, \gamma_x[j]} \left(\frac{1}{dx} \sum_{n=1}^N  a_n \left(v_x[i,j+n-\frac{1}{2},it+1] - v_x[i,j-n+\frac{1}{2},it+1] \right) \right) + dt \cdot f[k+1] 
\end{split}
\label{eq33}
\end{equation}
we create diagonal matrices ${\bf M}_{px}{\bf B}_{px}$ and ${\bf M}_{px}{\bf B}_{vx}$, both are size of $N_z \cdot N_x \times N_z \cdot N_x$,  which save the elements of pre-computed $\frac{2- dt\, \gamma_x[j]}{2+ dt\, \gamma_x[j]}$ and $\frac{2k[i,j] \, dt}{2+ dt\, \gamma_x[j]}$. With the help of these sparse matrices, the vectorized form of equation \ref{eq33} can be written as
\begin{equation}
{\bf p}_x^{it+\frac{3}{2}} =  {\bf M}_{px}{\bf B}_{px} \cdot {\bf p}_x^{it+\frac{1}{2}} +  {\bf M}_{px}{\bf B}_{vx} \cdot {\bf d}_v{\bf d}_x \cdot {\bf v}_x^{it+1} + dt \cdot {\bf f}^{it+1}
\label{eq34}
\end{equation}
Outside of the PML absorbing area, the damping coefficients $\gamma_x$ is reduced to $0$, so the equation \ref{eq34} can be simplified as 
\begin{equation}
\begin{split}
&p_x[i,j,it+\frac{3}{2}] =  p_x[i,j,it+\frac{1}{2}]  \\
&+k[i,j]\, dt \left(\frac{1}{dx} \sum_{n=1}^N  a_n \left(v_x[i,j+n-\frac{1}{2},it+1] - v_x[i,j-n+\frac{1}{2},it+1] \right) \right) + dt \cdot f[k+1]
\end{split}
\label{eq35}
\end{equation}
We also define sparse matrices ${\bf R}_{px}{\bf B}_{vx}$ and ${\bf r}_{v}{\bf d}_{x}$, so the vectorized form of equation \ref{eq35} can be written as
\begin{equation}
{\bf p}_x^{it+\frac{3}{2}} =  {\bf p}_x^{it+\frac{1}{2}} +  {\bf R}_{px}{\bf B}_{vx} \cdot {\bf r}_v{\bf d}_x \cdot {\bf v}_x^{it+1} + dt \cdot {\bf f}^{it+1}
\label{eq36}
\end{equation}
To facilitate the program, we also give the details of the partial differential operator as
\begin{equation}
\begin{bmatrix}
\frac{\partial v_x}{\partial x}[1] \\
\frac{\partial v_x}{\partial x}[2] \\
\frac{\partial v_x}{\partial x}[3] \\
\frac{\partial v_x}{\partial x}[4] \\
\frac{\partial v_x}{\partial x}[5] \\
\end{bmatrix}
= \frac{1}{dx} \begin{bmatrix}
a_1 & a_2 & & &\\
-a_1& a_1 & a_2 & & \\
-a_2& -a_1 & a_1 & a_2 & \\
& -a_2 & -a_1 & a_1 & a_2 \\
& & -a_2 & -a_1 & a_1 \\
\end{bmatrix}
\begin{bmatrix}
v_x[1+\frac{1}{2}] \\
v_x[2+\frac{1}{2}] \\
v_x[3+\frac{1}{2}] \\
v_x[4+\frac{1}{2}] \\
v_x[5+\frac{1}{2}] \\
\end{bmatrix}
\label{eq37}
\end{equation}
and its adjoint operator
\begin{equation}
\begin{bmatrix}
v_x[1+\frac{1}{2}] \\
v_x[2+\frac{1}{2}] \\
v_x[3+\frac{1}{2}] \\
v_x[4+\frac{1}{2}] \\
v_x[5+\frac{1}{2}] \\
\end{bmatrix}
= - \frac{1}{dx} \begin{bmatrix}
-a_1 & a_1 & a_2 & & \\
-a_2 & -a_1 & a_1 & a_2 &   \\
& -a_2 & -a_1 & a_1 & a_2  \\
& & -a_2 & -a_1 & a_1  \\
& & & -a_2 & -a_1  \\
\end{bmatrix}
\begin{bmatrix}
\frac{\partial v_x}{\partial x}[1] \\
\frac{\partial v_x}{\partial x}[2] \\
\frac{\partial v_x}{\partial x}[3] \\
\frac{\partial v_x}{\partial x}[4] \\
\frac{\partial v_x}{\partial x}[5] \\
\end{bmatrix}
\label{eq38}
\end{equation}

\subsection{Summarize the discretized wave equations}
The discretized wave-equation is given as
\begin{equation}
\begin{split}
& v_z[i+\frac{1}{2},j,it+1] =  \frac{2- dt\, \gamma_z[i+\frac{1}{2}]}{(2+ dt\, \gamma_z[i+\frac{1}{2}])}v_z[i+\frac{1}{2},j,it] \\
&+ \frac{2b[i+\frac{1}{2},j] \, dt}{(2+ dt\, \gamma_z[i+\frac{1}{2}])} \left(\frac{1}{dz} \sum_{n=1}^N  a_n \left( p[i+n,j,it+\frac{1}{2}] - p[i-n+1,j,it+\frac{1}{2}] \right)  \right) \\
& v_x[i,j+\frac{1}{2},it+1] =  \frac{2- dt\, \gamma_x[j+\frac{1}{2}]}{2+ dt\, \gamma_x[j+\frac{1}{2}]}v_x[i,j+\frac{1}{2},it] \\
& + \frac{2b[i,j+\frac{1}{2}] \, dt}{2+ dt\, \gamma_x[j+\frac{1}{2}]}
\left( \frac{1}{dx} \sum_{n=1}^N  a_n \left( p[i,j+n,it+\frac{1}{2}] - p[i,j-n+1,it+\frac{1}{2}]\right) \right) \\
& p_z[i,j,it+\frac{3}{2}] = \frac{2-dt \, \gamma_z[i]}{2+dt \, \gamma_z[i]} p_z[i,j,it+\frac{1}{2}]  \\
& + \frac{2k[i,j] \, dt}{2+dt \, \gamma_z[i] } \left(\frac{1}{dz} \sum_{n=1}^N  a_n \left( v_z[i+n-\frac{1}{2},j,it+1] - v_z[i-n+\frac{1}{2},j,it+1] \right)  \right) \\
&p_x[i,j,it+\frac{3}{2}] = \frac{2-dt \, \gamma_x[j]}{2+dt \, \gamma_x[j]} p_x[i,j,it+\frac{1}{2}]  \\
&+ \frac{2k[i,j]\, dt}{2+dt \, \gamma_x[j]} \left(\frac{1}{dx} \sum_{n=1}^N  a_n \left(v_x[i,j+n-\frac{1}{2},it+1] - v_x[i,j-n+\frac{1}{2},it+1] \right) \right) + dt \cdot f[k+1]  \\
\end{split}
\label{eq39}
\end{equation}
These equation can be written as the sparse matrix-vector format as
\begin{equation}
\begin{split}
{\bf v}_z^{it+1} &= {\bf M}_{vz}{\bf B}_{vz}  \cdot {\bf v}_z^{it} +  {\bf M}_{vz}{\bf B}_{p} \cdot {\bf d}_p{\bf d}_z \cdot {\bf p}^{it+\frac{1}{2}} \\
{\bf v}_x^{it+1} &=  {\bf M}_{vx}{\bf B}_{vx} \cdot {\bf v}_x^{it} +  {\bf M}_{vx}{\bf B}_{p} \cdot {\bf d}_p{\bf d}_x \cdot {\bf p}^{it+\frac{1}{2}}  \\
{\bf p}_z^{it+\frac{3}{2}} &= {\bf M}_{pz}{\bf B}_{pz}  \cdot {\bf p}_z^{it+\frac{1}{2}} +  {\bf M}_{pz}{\bf B}_{vz} \cdot {\bf d}_v{\bf d}_z \cdot {\bf v}_z^{it+1}  \\
{\bf p}_x^{it+\frac{3}{2}} &=  {\bf M}_{px}{\bf B}_{px} \cdot {\bf p}_x^{it+\frac{1}{2}} +  {\bf M}_{px}{\bf B}_{vx} \cdot {\bf d}_v{\bf d}_x \cdot {\bf v}_x^{it+1} + dt \cdot {\bf f}^{it+1}  \\
\end{split}
\label{eq40}
\end{equation}
We define the the state variables, which is obtained by vertical concatenate the wave field, at two neighbouring time steps as
\begin{equation}
{\bf u}^{it} = \begin{bmatrix}
{\bf v}_z^{it} \\
{\bf v}_x^{it} \\
{\bf p}_z^{it+\frac{1}{2}} \\
{\bf p}_x^{it+\frac{1}{2}} \\
\end{bmatrix}, \,\,\, 
{\bf u}^{it+1} = \begin{bmatrix}
{\bf v}_z^{it+1} \\
{\bf v}_x^{it+1} \\
{\bf p}_z^{it+\frac{3}{2}} \\
{\bf p}_x^{it+\frac{3}{2}} \\
\end{bmatrix},
\label{eq41}
\end{equation}

@@

&+ dt \cdot f[k+1]   

@@

\frac{\partial {\bf L}}{\partial m_l} {\bf u}_{k-1} &=  

>>>>

We use stagger grid finite difference method to discretize equation \ref{eq2}, the particle $v_z$ is put on the stagger grid $[i+\frac{1}{2},j,it]$, horizontal particle velocity $v_x$ is put on the stagger grid $[i, j+\frac{1}{2}, it]$, and the pressure component is put on the grid $[i,j,it+\frac{1}{2}]$. Where $i,j,it$ are the grid index in vertical, horizontal, time directions, respectively. Figure \ref{f1} shows the locations of wave field components and model parameters, respectively.
\begin{figure}[htb] 
   \centering
   \includegraphics[width=0.6\columnwidth]{Fig/f1.pdf} 
   \caption{Stagger grid}
   \label{f1}
\end{figure}

@@

\subsection{${\bf p}_x$ time-stepping}
\begin{equation}
\frac{\partial p_x}{\partial t} + \gamma_x p_x = k \frac{\partial v_x}{\partial x} + f(t) 
\label{eq32}
\end{equation}
The finite difference form of this equation is given as
\begin{equation}
\begin{split}
&p_x[i,j,it+\frac{3}{2}] = \frac{2-dt \, \gamma_x[j]}{2+dt \, \gamma_x[j]} p_x[i,j,it+\frac{1}{2}]  \\
&+ \frac{2k[i,j]\, dt}{2+dt \, \gamma_x[j]} \left(\frac{1}{dx} \sum_{n=1}^N  a_n \left(v_x[i,j+n-\frac{1}{2},it+1] - v_x[i,j-n+\frac{1}{2},it+1] \right) \right) + dt \cdot f[it+1] 
\end{split}
\label{eq33}
\end{equation}
we create diagonal matrices ${\bf M}_{px}{\bf B}_{px}$ and ${\bf M}_{px}{\bf B}_{vx}$, both are size of $N_z \cdot N_x \times N_z \cdot N_x$,  which save the elements of pre-computed $\frac{2- dt\, \gamma_x[j]}{2+ dt\, \gamma_x[j]}$ and $\frac{2k[i,j] \, dt}{2+ dt\, \gamma_x[j]}$. With the help of these sparse matrices, the vectorized form of equation \ref{eq33} can be written as
\begin{equation}
{\bf p}_x^{it+\frac{3}{2}} =  {\bf M}_{px}{\bf B}_{px} \cdot {\bf p}_x^{it+\frac{1}{2}} +  {\bf M}_{px}{\bf B}_{vx} \cdot {\bf d}_v{\bf d}_x \cdot {\bf v}_x^{it+1} + dt \cdot {\bf f}^{it+1}
\label{eq34}
\end{equation}
Outside of the PML absorbing area, the damping coefficients $\gamma_x$ is reduced to $0$, so the equation \ref{eq34} can be simplified as 
\begin{equation}
\begin{split}
&p_x[i,j,it+\frac{3}{2}] =  p_x[i,j,it+\frac{1}{2}]  \\
&+k[i,j]\, dt \left(\frac{1}{dx} \sum_{n=1}^N  a_n \left(v_x[i,j+n-\frac{1}{2},it+1] - v_x[i,j-n+\frac{1}{2},it+1] \right) \right) + dt \cdot f[it+1]
\end{split}
\label{eq35}
\end{equation}
We also define sparse matrices ${\bf R}_{px}{\bf B}_{vx}$ and ${\bf r}_{v}{\bf d}_{x}$, so the vectorized form of equation \ref{eq35} can be written as
\begin{equation}
{\bf p}_x^{it+\frac{3}{2}} =  {\bf p}_x^{it+\frac{1}{2}} +  {\bf R}_{px}{\bf B}_{vx} \cdot {\bf r}_v{\bf d}_x \cdot {\bf v}_x^{it+1} + dt \cdot {\bf f}^{it+1}
\label{eq36}
\end{equation}
To facilitate the program, we also give the details of the partial differential operator as
\begin{equation}
\begin{bmatrix}
\frac{\partial v_x}{\partial x}[1] \\
\frac{\partial v_x}{\partial x}[2] \\
\frac{\partial v_x}{\partial x}[3] \\
\frac{\partial v_x}{\partial x}[4] \\
\frac{\partial v_x}{\partial x}[5] \\
\end{bmatrix}
= \frac{1}{dx} \begin{bmatrix}
a_1 & a_2 & & &\\
-a_1& a_1 & a_2 & & \\
-a_2& -a_1 & a_1 & a_2 & \\
& -a_2 & -a_1 & a_1 & a_2 \\
& & -a_2 & -a_1 & a_1 \\
\end{bmatrix}
\begin{bmatrix}
v_x[1+\frac{1}{2}] \\
v_x[2+\frac{1}{2}] \\
v_x[3+\frac{1}{2}] \\
v_x[4+\frac{1}{2}] \\
v_x[5+\frac{1}{2}] \\
\end{bmatrix}
\label{eq37}
\end{equation}
and its adjoint operator
\begin{equation}
\begin{bmatrix}
v_x[1+\frac{1}{2}] \\
v_x[2+\frac{1}{2}] \\
v_x[3+\frac{1}{2}] \\
v_x[4+\frac{1}{2}] \\
v_x[5+\frac{1}{2}] \\
\end{bmatrix}
= - \frac{1}{dx} \begin{bmatrix}
-a_1 & a_1 & a_2 & & \\
-a_2 & -a_1 & a_1 & a_2 &   \\
& -a_2 & -a_1 & a_1 & a_2  \\
& & -a_2 & -a_1 & a_1  \\
& & & -a_2 & -a_1  \\
\end{bmatrix}
\begin{bmatrix}
\frac{\partial v_x}{\partial x}[1] \\
\frac{\partial v_x}{\partial x}[2] \\
\frac{\partial v_x}{\partial x}[3] \\
\frac{\partial v_x}{\partial x}[4] \\
\frac{\partial v_x}{\partial x}[5] \\
\end{bmatrix}
\label{eq38}
\end{equation}

\subsection{Summarize the discretized wave equations}
The discretized wave-equation is given as
\begin{equation}
\begin{split}
& v_z[i+\frac{1}{2},j,it+1] =  \frac{2- dt\, \gamma_z[i+\frac{1}{2}]}{(2+ dt\, \gamma_z[i+\frac{1}{2}])}v_z[i+\frac{1}{2},j,it] \\
&+ \frac{2b[i+\frac{1}{2},j] \, dt}{(2+ dt\, \gamma_z[i+\frac{1}{2}])} \left(\frac{1}{dz} \sum_{n=1}^N  a_n \left( p[i+n,j,it+\frac{1}{2}] - p[i-n+1,j,it+\frac{1}{2}] \right)  \right) \\
& v_x[i,j+\frac{1}{2},it+1] =  \frac{2- dt\, \gamma_x[j+\frac{1}{2}]}{2+ dt\, \gamma_x[j+\frac{1}{2}]}v_x[i,j+\frac{1}{2},it] \\
& + \frac{2b[i,j+\frac{1}{2}] \, dt}{2+ dt\, \gamma_x[j+\frac{1}{2}]}
\left( \frac{1}{dx} \sum_{n=1}^N  a_n \left( p[i,j+n,it+\frac{1}{2}] - p[i,j-n+1,it+\frac{1}{2}]\right) \right) \\
& p_z[i,j,it+\frac{3}{2}] = \frac{2-dt \, \gamma_z[i]}{2+dt \, \gamma_z[i]} p_z[i,j,it+\frac{1}{2}]  \\
& + \frac{2k[i,j] \, dt}{2+dt \, \gamma_z[i] } \left(\frac{1}{dz} \sum_{n=1}^N  a_n \left( v_z[i+n-\frac{1}{2},j,it+1] - v_z[i-n+\frac{1}{2},j,it+1] \right)  \right) \\
&p_x[i,j,it+\frac{3}{2}] = \frac{2-dt \, \gamma_x[j]}{2+dt \, \gamma_x[j]} p_x[i,j,it+\frac{1}{2}]  \\
&+ \frac{2k[i,j]\, dt}{2+dt \, \gamma_x[j]} \left(\frac{1}{dx} \sum_{n=1}^N  a_n \left(v_x[i,j+n-\frac{1}{2},it+1] - v_x[i,j-n+\frac{1}{2},it+1] \right) \right) + dt \cdot f[it+1]  \\
\end{split}
\label{eq39}
\end{equation}
These equation can be written as the sparse matrix-vector format as
\begin{equation}
\begin{split}
{\bf v}_z^{it+1} &= {\bf M}_{vz}{\bf B}_{vz}  \cdot {\bf v}_z^{it} +  {\bf M}_{vz}{\bf B}_{p} \cdot {\bf d}_p{\bf d}_z \cdot {\bf p}^{it+\frac{1}{2}} \\
{\bf v}_x^{it+1} &=  {\bf M}_{vx}{\bf B}_{vx} \cdot {\bf v}_x^{it} +  {\bf M}_{vx}{\bf B}_{p} \cdot {\bf d}_p{\bf d}_x \cdot {\bf p}^{it+\frac{1}{2}}  \\
{\bf p}_z^{it+\frac{3}{2}} &= {\bf M}_{pz}{\bf B}_{pz}  \cdot {\bf p}_z^{it+\frac{1}{2}} +  {\bf M}_{pz}{\bf B}_{vz} \cdot {\bf d}_v{\bf d}_z \cdot {\bf v}_z^{it+1}  \\
{\bf p}_x^{it+\frac{3}{2}} &=  {\bf M}_{px}{\bf B}_{px} \cdot {\bf p}_x^{it+\frac{1}{2}} +  {\bf M}_{px}{\bf B}_{vx} \cdot {\bf d}_v{\bf d}_x \cdot {\bf v}_x^{it+1} + dt \cdot {\bf f}^{it+1}  \\
\end{split}
\label{eq40}
\end{equation}
We define the the state variables, which is obtained by vertical concatenate the wave field, at two neighbouring time steps as
\begin{equation}
{\bf u}^{it} = \begin{bmatrix}
{\bf v}_z^{it} \\
{\bf v}_x^{it} \\
{\bf p}_z^{it+\frac{1}{2}} \\
{\bf p}_x^{it+\frac{1}{2}} \\
\end{bmatrix}, \,\,\, 
{\bf u}^{it+1} = \begin{bmatrix}
{\bf v}_z^{it+1} \\
{\bf v}_x^{it+1} \\
{\bf p}_z^{it+\frac{3}{2}} \\
{\bf p}_x^{it+\frac{3}{2}} \\
\end{bmatrix},
\label{eq41}
\end{equation}

@@

&+ dt \cdot f[it+1]   

@@

\frac{\partial {\bf L}}{\partial m_l} {\bf u}_{it-1} &=  
